Applications of Even-point Muiltiloop Unitarity: Exponentiation, Anomalies and Evanescence

### 2023-05-29 eventPointTwoLoop.tex
nameChange and sketch review/method rewrite

\newcommand{\afour}{\ensuremath{A_4^{\text{tree}}}}
\definecolor{NUpurple}{RGB}{078,042,132}

@@

\title{Exponentiation, anomalies and evanescence beyond one-loop}

\abstract{\dj{why initiate?}
In this work we study multi-loop amplitudes and quantum actions for a special class of effective field theories which includes both the nonlinear sigma model (NLSM) and its double-copy gauge theories such as Born-Infeld and supersymmetric generalizations. In computing these amplitudes through two-loops, we identify a wide variety of rich physical structure. First, we find that the leading IR divergence of NLSM amplitudes exponentiates. Second, we find that the counterterms required to restore U(1) behavior at loop-level in the Born-Infeld theory can be constructed via a symmetric-structure double-copy.  We demonstrate that the one-minus $(-+++)$ two-loop amplitude vanishes upon the introduction of an evanescent operator in the $D$-dimensional description of Born-Infeld theory. In addition to these pure photon counterterms, we show through explicit calculation that the anomalous matrix elements that violates $U(1)$ duality invariance can be similarly cancelled by summing over internal $\mathcal{N}=4$ DBIVA superfields. In doing so, we find that $\mathcal{N}=4$ Dirac-Born-Infeld-Volkov-Akulov (DBIVA) amplitudes permit double-copy construction through two-loop order. This is shown by reproducing our unitarity based result with a loop-level double copy of $\mathcal{N}=4$ super-Yang-Mills numerators and our two-loop NLSM amplitudes. This result provides strong evidence for the existence of a color-dual representations for NLSM beyond one-loop. We conclude with an overview of how $D$-dimensional four-photon counterterms can be constructed in generality with the symmetric-structure double-copy, and outline a convenient way of counting evanescent operators at higher orders using Hilbert series as generating functions. }

@@

\section{Introduction}\label{sec:intro}
Here we study  effective field theory (EFT) at the multi-loop level using a combination of generalized unitarity \cite{UnitarityMethod, Fusing, BDKUniarityReview} and the double-copy construction \cite{BCJ, BCJLoop}. The EFTs that we consider in this work are the nonlinear sigma model (NLSM) and a related family of gauge theories involving Born-Infeld theory and supersymmetric generalizations known as Dirac-Born-Infeld-Volkov-Akulov (DBIVA) theories. These families are known at tree-level to be double-copies between NLSM and Yang-Mills theories with varying amounts of supersymmetry.

The motivation for this work is three-fold. First, perturbative calculations in NLSM and DBIVA effective field theories are significantly simpler than the typical multi-loop calculations of renormalizable theories, like quantum-chromodynamics (QCD) and super-Yang-Mills (sYM). As we will show, the multi-loop amplitudes for these theories allow us to recycle many of the $D$-dimensional integration tools that are so powerful at one-loop order. This allows us to compute a large catalog of $D$-dimensional two-loop amplitudes, which provides fertile ground for cultivating insights about the multi-loop structure of effective field theories. The perturbative depth of our calculations sheds light on the exponential structure of IR divergences for NLSM amplitudes through two-loops, the anomalous behavior $U(1)$ duality invariance through two-loop order in DBIVA theories, as well as the emergence of evanescent operators relevant for anomaly cancellation in pure Born-Infeld theory. We believe these surprisingly rich physical structures could serve as a theoretical laboratory for future studies of the interplay between effective field theory operators and perturbative calculations in quantum field theory.  \dj{Maybe we get rid of numbers.  we say three-fold motivation, then give three enumerated things, but that's only in the first motivation.}

@@

Lastly, DBIVA effective field theories touch a wide range research areas at the forefront of high energy physics. Of more formal interest, is the presence of quantum anomalies that violate the $U(1)$ duality invariance, which the theory enjoys at tree-level \cite{Elvang:2020kuj}. The presence of these $U(1)$ anomalies in gravity is closely linked to ultraviolet divergences computed at loop-level \cite{Bern:2007xj,Carrasco:2013ypa,Craig:2019zkf,Monteiro:2022nqt}. It has been argued that cancelling these anomalous matrix elements with $R^n$ counterterms can lead to enhanced UV cancellations \cite{Bern:2017tuc,Bern:2017rjw,Bern:2019isl}. This relationship has been demonstrated both for pure Einstein-Hilbert gravity \cite{Goroff:1985sz}, and also for less than maximal $\mathcal{N}\leq 4$ supergravity \cite{Bern:2013uka}. While DBIVA theories are themselves are ultraviolet divergent in $D=4$, due to their simplicity they serve as an essential laboratory for probing anomaly cancellation at high loop order. Moreover, in addition to their formal theory relevance, DBI has garnered wide phenomenological interest in cosmology, both for sourcing non-gaussianities in CMB bispectrum \cite{Alishahiha:2004eh,Creminelli:2005hu,Fergusson:2008ra}, and for their consistency with the observed CMB tensor mode suppression \cite{Carrasco:2015pla,Carrasco:2015rva,Carrasco:2015uma,BICEP:2021xfz,Kallosh:2021mnu}. Thus, with inflationary data on the horizon, understanding the perturbative structure of DBI could be particularly relevant for modeling early universe quantum fluctuations.

Our paper is organized as follows. To begin, we provide a brief review of the types of models studied in this work, namely NLSM and DBIVA. While NLSM in general carries nonabelian color structure, we call such theories \textit{abelianized} because the both appear naturally in the field theory limit of the \textit{abelianized open superstring} \cite{Green:1982sw}. Tree-level amplitudes for the open superstring (OSS) ~\cite{Mafra:2011nv,Mafra:2011nw}  can be constructed from a double copy \cite{Broedel:2013tta} between $\mathcal{N}=N_f$ super-Yang-Mills (sYM) and $Z$-theory amplitudes~\cite{Carrasco:2016ldy,Carrasco:2016ygv,Mafra:2016mcc}, as follows,
\begin{equation}
A^{\text{OSS}}= Z \otimes A^{ \text{sYM}}
\end{equation}
where $\otimes$ implements the field theory double copy \cite{KLT, BCJ}. To recover the abelian sector, one simply sums over all color orderings on the side of the bi-colored $Z$-theory amplitudes that obey string monodromy relations. In the field theory limit, the observables for so-called \textit{abelianized} $Z$-theory are simply the amplitudes generated by the NLSM Lagrangian \cite{Carrasco:2016ldy}. That is, given a $n$-point amplitude in $Z$-theory, with field theory ordering $a=(a_1,a_2,...,a_n)$ and string theoretic color ordering $A=(A_1,A_2,...,A_n)$, one finds the following relation:
\begin{equation}
A^{\text{NLSM}}{(a_1,a_2,...,a_n)} \equiv \lim_{\alpha' \to 0} (\alpha')^{2-n}  Z_{\times}{(a_1,a_2,...,a_n)}
\end{equation}
where $Z_{\times}$ are abelianized $Z$-theory amplitudes of \cite{Carrasco:2016ldy}, defined by summing all possible orderings of Chan-Paton factors:
\begin{equation}
Z_{\times}(a_1,a_2,...,a_n)\equiv \sum_{A \in S^{n-1}} Z_{(A_1,...,A_{n-1},n)}{(a_1,a_2,...,a_n)}
\end{equation}
Since the field theory limit of abelian $Z$-theory produces NLSM amplitudes, the field theory limit of the abelianized OSS, gives rise to DBIVA observables at leading order:
\begin{equation}
\boxed
{\begin{aligned} 
\lim_{\alpha' \to 0 }Z^{\text{OSS}}_{\times}&= A^{\text{NLSM}} \otimes A^{\text{BAS}}\equiv A^{\text{NLSM}}
\\
\Rightarrow \quad\lim_{\alpha' \to 0 }A^{\text{OSS}}_{\times}&= A^{\text{NLSM}} \otimes A^{ \text{sYM}}\equiv A^{\text{DBIVA}}
\end{aligned}}
\end{equation}
These two classes of amplitudes, NLSM and DBIVA, are the observables that we will consider in the multi-loop studies of this work. Additional details on the bi-colored $Z$-theory amplitudes these string-theoretic double-copy constructions can be found in \cite{Broedel:2013tta,Carrasco:2016ldy,Carrasco:2016ygv,Mafra:2016mcc,Azevedo:2018dgo} and references therein.

By far the most important feature of these amplitudes relevant for our analysis is that they are generated by purely \textit{even-point} contacts. That is, the effective actions of abelianized theories are consistent with the following interaction Lagrangian:
\begin{equation}
\mathcal{L}^{\text{abelian}}_{\text{int.}} = \sum_{k=2} \alpha'^{n-2} \mathcal{O}^{(2n)}_{2k} 
\end{equation}
where $2k$ counts the field content of the effective operator, and $2n$ counts the mass-dimension determined by the number of fields and derivatives contained in the operator. The dimensionful scale, $\alpha'$, carries mass dimension $[\alpha']=-2$ in $D=4$.

The even-point nature of this class of theories make them comparatively easy to compute relative to their arbitrary-multiplicity cousins. Indeed, the observation that will drive many of the multi-loop results is the following: \textit{amplitudes for abelianized theories at next-to-next-to-leading-order (NNLO) in $\alpha'$ are recursively one-loop}. To gain some familiarity with why this is the case, consider the interactions needed to construct two loop amplitudes for a simple $\varphi^{2k}$-scalar theory:
\begin{equation}\label{eq:evenPointL}
\mathcal{L}_{\varphi^{2k}} = \frac{1}{2}(\partial \varphi)^2 + c_4 \varphi^4 + c_6 \varphi^6+ c_8 \varphi^8+\cdots 
\end{equation}
While the Feynman rules for this theory are trivial, unlike those needed for the vectors and fermions in the DBIVA supermultiplet and the derivatively coupled pions in NLSM, the Feynman diagrams needed to compute amplitudes are identical to those of $A^{\text{OSS}}_{\times}$. Indeed, the two-loop amplitude for $\varphi^{2k}$ theory is generated simply by evaluating the following set of scalar 1PI integrals:

@@

The outline of the paper is as follows: in \sect{sec:IntMethods}, we will provide an overview of the on-shell methods and integration methods needed to probe the multi-loop physics studied in this paper. In \sect{sec:Loops}, we present our results, beginning with a warm-up calculation in \sect{sec:NLSMU} where we construct the integrands for NLSM through NNLO in the effective coupling, and compute the fully integrated amplitudes. We show that in $D=2-2\epsilon$, the leading IR divergences of the theory exponentiate. With the scaffolding of $D$-dimensional integration in hand, we then compute two-loop amplitudes for $\mathcal{N}=4$ DBIVA theory and pure-photon Born-Infeld theory in \sect{sec:DBIU}. There we compute the anomalies present beyond one-loop order.  Then, using the NLSM integrands of \sect{sec:NLSMU}, we perform a multi-loop double-copy to $\mathcal{N}=4$ DBIVA observables with the color-dual basis numerators available in the literature in \sect{sec:DBIvDC}. After computing the catalog of two-loop amplitudes, we study the construction of quantum effective actions in \sect{sec:Actions} as they relate to the anomalous matrix elements present in two-loop Born-Infeld amplitudes. We demonstrate in \sect{sec:Anomalies} that the one-minus anomaly at two-loop requires the introduction of an evanescent operator at $\mathcal{O}(\alpha'^4)$. Then in \sect{sec:ActionDC} we take the opportunity to discuss the application of symmetric-structure double copy to cancel these anomalies, along with their relationship to higher-spin modes as studied in a recent work by the authors. Finally, in \sect{sec:EOpsCounting} we lay out a Hilbert series framework for counting evanescent operators at general orders in mass-dimension. To conclude, in \sect{sec:Discussion} we discuss many directions of future work and summarize the insights gained from this study. 
\section{Covariant Integration Methods}\label{sec:IntMethods}
\subsection{Generalized Unitarity}\label{sec:genU}
To perform the loop-level calculations in this paper, we will construct the integrand in $D$-dimensions using generalized unitarity cuts \cite{Bern:1995db,Bern:1996ja}. In the kinematic regime where the internal loop momenta go on-shell, the residue is a product of on-shell tree level diagrams, summed over all physical states that can cross the cut. For example, the kinematic numerator of the one-loop integrand $\mathcal{I}_{\text{1-loop}}$ of pure Born-Infeld theory should satisfy the following constraint:
\label{GUCut}
\text{Cut}( \mathcal{I}_{\text{1-loop}}) = \sum_{h_i\in \text{states}} \mathcal{M}^{\text{BI}}_4(l_1^{h_1},-l_2^{\bar{h}_2})\mathcal{M}^{\text{BI}}_4(l_2^{h_2},-l_1^{\bar{h}_1})\,,
where $\text{Cut}( \mathcal{I}_{\text{1-loop}}) $ extracts the kinematic numerator by imposing a maximal cut on the one-loop integrand 
\text{Cut}( \mathcal{I}_{\text{1-loop}})  \equiv l_1^2l_2^2(\mathcal{I}_{\text{1-loop}})\Big|_{l_1^2,l_2^2\rightarrow 0}\,.
To carry out the sum over states in \eqn{GUCut} we apply the following $D$-dimensional completeness relation for on-shell polarizations,
\begin{equation}\label{stateSum}
\sum_{\text{states}}\varepsilon^\mu_{(l)}\varepsilon^\nu_{(-l)} = \eta^{\mu\nu} - \frac{l^\mu q^\nu+l^\nu q^\mu}{l\cdot q},
with null reference momentum, $q^2=0$. The dependence of $q$ is a gauge choice that disappears when on-shell kinematic constraints are imposed. At loop level, the state-sum of \eqn{stateSum} can potentially lead to dimension dependence in the $D$-dimensional integrand. This is due to terms that contain products of internal polarizations:
 \mathcal{I}_{\text{1-loop}} &\supset \sum_{ \text{states}} \varepsilon^\mu_{(l)}\varepsilon^\nu_{(-l)} \eta_{\mu\nu}
 &=\left(\eta^{\mu\nu} - \frac{l^\mu q^\nu+l^\nu q^\mu}{l\cdot q}\right) \eta_{\mu\nu}
 &=D_s-2
 where $D_s$ is the spin-dimension of the internal photons, and where $D_s-2$ just counts the number of photon states. To calculate the loop-level amplitudes in this work, we use dimensional regularization to regulate the integrals. In this work we adopt the conventions of  \cite{collins_1984,Bern:2002zk} and take the spin-dimension, $D_s$, to be
the same as the dimension appearing in dimensional regularization, $D=4-2\epsilon$. We will use them interchangeably throughout. As shown in \eqn{eq:2loopEven} of the introduction, many of the diagrams that contribute to abelianized loop-level amplitudes are scaleless, and thus vanish in dimensional regularization.

>>>>

\newcommand{\afour}{\ensuremath{A_4^{\text{tree}}}}

\definecolor{NUpurple}{RGB}{078,042,132}

@@

\title{Applications of Even-point Muiltiloop Unitarity: Exponentiation, Anomalies and Evanescence}

\abstract{
We identify novel  structure in newly computed multi-loop amplitudes and quantum actions for a special class of effective field theories which includes both the nonlinear sigma model (NLSM) and its double-copy gauge theories such as Born-Infeld and supersymmetric generalizations.  We exploit special properties of these even-point theories to generalize the direct extraction of Feynman integral coefficients using $D$-dimensional unitarity cuts to the multi-loop case for the first time.  Doing so, we find that the leading IR divergence of NLSM amplitudes exponentiates. Second, we find that the counterterms required to restore U(1) behavior at loop-level in the Born-Infeld theory can be constructed via a symmetric-structure double-copy.  We demonstrate that the one-minus $(-+++)$ two-loop amplitude vanishes upon the introduction of an evanescent operator in the $D$-dimensional description of Born-Infeld theory. In addition to these pure photon counterterms, we verify through explicit calculation that the anomalous matrix elements that violates $U(1)$ duality invariance can be alternatvely cancelled by summing over internal $\mathcal{N}=4$ DBIVA superfields. Finally we find that $\mathcal{N}=4$ Dirac-Born-Infeld-Volkov-Akulov (DBIVA) amplitudes permit double-copy construction through two-loop order. This is shown by reproducing our unitarity based result with a loop-level double copy between $\mathcal{N}=4$ super-Yang-Mills numerators and our two-loop NLSM amplitudes. This result provides strong evidence for the existence of a color-dual representations for NLSM beyond one-loop. We conclude with an overview of how $D$-dimensional four-photon counterterms can be constructed in generality with the symmetric-structure double-copy, and outline a convenient way of counting evanescent operators at higher orders using Hilbert series as generating functions. }

@@

\section{Introduction}\label{sec:intro}
\dj{Probably a more basic introduction for people who don't know what generalized unitarity is, or double copy, poor souls.}
\dj{Emphasize one advantage of unitarity methods at 1-loop is the ability to integrate by doing cuts -- i.e. know exactly the coefficients of each basis integral.  We show for a very special class of theories we can morally do exactly the same at higher loops -- why? even points,recursion blah blah balh.  In sect 2 we review all the old.  In sect 3 we lay out Even-Point Multiloop Unitarity in detail what properties we rely on and how we exploit them blah blah.  }

Here we study  effective field theory (EFT) at the multi-loop level using a combination of generalized unitarity \cite{UnitarityMethod, Fusing, BDKUniarityReview} and the double-copy construction \cite{BCJ, BCJLoop}. The EFTs that we consider in this work are the nonlinear sigma model (NLSM) and a related family of gauge theories involving Born-Infeld theory and supersymmetric generalizations known as Dirac-Born-Infeld-Volkov-Akulov (DBIVA) theories. These families are known at tree-level to be double-copies between NLSM and Yang-Mills theories with varying amounts of supersymmetry.

The motivation for this work is three-fold. First, perturbative calculations in NLSM and DBIVA effective field theories are significantly simpler than the typical multi-loop calculations of renormalizable theories, like quantum-chromodynamics (QCD) and super-Yang-Mills (sYM). As we will show, the multi-loop amplitudes for these theories allow us to recycle many of the $D$-dimensional integration tools that are so powerful at one-loop order. This allows us to compute a large catalog of $D$-dimensional two-loop amplitudes, which provides fertile ground for cultivating insights about the multi-loop structure of effective field theories. The perturbative depth of our calculations sheds light on the exponential structure of IR divergences for NLSM amplitudes through two-loops, the anomalous behavior $U(1)$ duality invariance through two-loop order in DBIVA theories, as well as the emergence of evanescent operators relevant for anomaly cancellation in pure Born-Infeld theory. We believe these surprisingly rich physical structures could serve as a theoretical laboratory for future studies of the interplay between effective field theory operators and perturbative calculations in quantum field theory.

@@

Finally, DBIVA effective field theories touch a wide range research areas at the forefront of high energy physics. Of more formal interest, is the presence of quantum anomalies that violate the $U(1)$ duality invariance, which the theory enjoys at tree-level \cite{Elvang:2020kuj}. The presence of these $U(1)$ anomalies in gravity is closely linked to ultraviolet divergences computed at loop-level \cite{Bern:2007xj,Carrasco:2013ypa,Craig:2019zkf,Monteiro:2022nqt}. It has been argued that cancelling these anomalous matrix elements with $R^n$ counterterms can lead to enhanced UV cancellations \cite{Bern:2017tuc,Bern:2017rjw,Bern:2019isl}. This relationship has been demonstrated both for pure Einstein-Hilbert gravity \cite{Goroff:1985sz}, and also for less than maximal $\mathcal{N}\leq 4$ supergravity \cite{Bern:2013uka}. While DBIVA theories are themselves are ultraviolet divergent in $D=4$, due to their simplicity they serve as an essential laboratory for probing anomaly cancellation at high loop order. Moreover, in addition to their formal theory relevance, DBI has garnered wide phenomenological interest in cosmology, both for sourcing non-gaussianities in CMB bispectrum \cite{Alishahiha:2004eh,Creminelli:2005hu,Fergusson:2008ra}, and for their consistency with the observed CMB tensor mode suppression \cite{Carrasco:2015pla,Carrasco:2015rva,Carrasco:2015uma,BICEP:2021xfz,Kallosh:2021mnu}. Thus, with inflationary data on the horizon, understanding the perturbative structure of DBI could be particularly relevant for modeling early universe quantum fluctuations.

To approach these problems we generalize the by now standard one-loop unitarity approach of Forde to higher loops exploiting a special feature of our theories blah blah.

The outline of the paper is as follows: in \sect{sec:Review}, we will provide an overview of the on-shell methods and integration methods needed to probe the multi-loop physics studied in this paper, we also introduce the class of theories theories we will study. In \sect{sec:Loops}, we present our results, beginning with a warm-up calculation in \sect{sec:NLSMU} where we construct the integrands for NLSM through NNLO in the effective coupling, and compute the fully integrated amplitudes. We show that in $D=2-2\epsilon$, the leading IR divergences of the theory exponentiate. With the scaffolding of $D$-dimensional integration in hand, we then compute two-loop amplitudes for $\mathcal{N}=4$ DBIVA theory and pure-photon Born-Infeld theory in \sect{sec:DBIU}. There we compute the anomalies present beyond one-loop order.  Then, using the NLSM integrands of \sect{sec:NLSMU}, we perform a multi-loop double-copy to $\mathcal{N}=4$ DBIVA observables with the color-dual basis numerators available in the literature in \sect{sec:DBIvDC}. After computing the catalog of two-loop amplitudes, we study the construction of quantum effective actions in \sect{sec:Actions} as they relate to the anomalous matrix elements present in two-loop Born-Infeld amplitudes. We demonstrate in \sect{sec:Anomalies} that the one-minus anomaly at two-loop requires the introduction of an evanescent operator at $\mathcal{O}(\alpha'^4)$. Then in \sect{sec:ActionDC} we take the opportunity to discuss the application of symmetric-structure double copy to cancel these anomalies, along with their relationship to higher-spin modes as studied in a recent work by the authors. Finally, in \sect{sec:EOpsCounting} we lay out a Hilbert series framework for counting evanescent operators at general orders in mass-dimension. To conclude, in \sect{sec:Discussion} we discuss many directions of future work and summarize the insights gained from this study.

\section{Review} 
\label{sec:Review}
\subsection{color-dressed and ordered amplitudes}
\label{subsec:AmpReview}
\dj{2.1 amplitudes and nomenclature (integrand, numerators, denominators, symmetry factors, color-ordered amplitudes)., e.g. when we use a curly A or M what does it mean and why. Also throughout the paper we should standardize, $\{\cal{O},{\cal{M}}\}\to\cal{A}$ .  I guess we use straight $\epsilon$ for dimensions, and $\varepsilon$ for polarizations.  We can explain polarizations in a 4D vs D-dim part. I would do standard choices so normal people (like us) can skip this section but if they hand it to some rando QCD person or string person to referee they can figure out what we're talking about. I don't know, between a page or two here maybe.}

\subsection{Color-Kinematics Duality and Double-Copy Construction}
\label{subsec:DCReview}

\dj{color-kinematics and double-copy (this could literally be the standard paragraph or so, trivalent graphs, color-weights, maybe a little longer than usual because we're also introducing $d's$)}

\subsection{4D Spinor Helicity vs $D$-dimensions}
\label{subsec:4DandDDReview}
\dj{4D (spinor helicity) vs arb D dimensions and  formal polarizations. (maybe half a page).  }

\subsection{Even-Point Effective Field Theories}
\label{subsec:EPEFTReview}
\dj{EP theories and relevant effective actions, hilbert series, state of the art of what has been calculated in each theory. (move some of the introduction discussion here, but this could be a couple of pages.)}

We here provide a brief review of the types of models studied in this work, namely  NLSM and DBIVA.

\dj{copied from later}
The Lagrangian for the \textit{nonlinear sigma model} (NLSM) can be expressed in terms of the chiral current, $j_\mu = U^\dagger \partial_\mu U$, as follows,
\be
\mathcal{L}^{\text{NLSM}}= \frac{1}{2}\text{tr}[(\partial_\mu U)^\dagger (\partial ^\mu U)] = \frac{1}{2}\text{tr}\left[\frac{(\partial_\mu\pi )( \partial^\mu \pi) }{(1-f_\pi^{-2}\pi^2)^2}\right]
\ee
where the trace is over the color indices of the gauge group, $\pi \equiv \pi^a T^a$, and on the right hand side we have applied the Cayley parameterization to the $SU(N_c)$ group elements, $U= \frac{1+\pi/f_\pi}{1-\pi/f_\pi}$, where $f_\pi$ is the dimensionful pion decay width. As noted in the preamble to this section, to construct the integrand, we will impose a spanning set of generalized unitarity cuts that appear in a minimal color basis. This procedure will provide the scaffolding for the colorless amplitudes of interest in the latter part of the paper.  This lagrangian is the same independent of what dimension.

In contrast, maximally supersymmetric Born-Infeld comes from adding the maximal amount of supersymmetric fermionic and scalar matter to Born-Infeld.  The Born-Infeld Lagrangian, which describes the pure vector sector of DBIVA theories, can be expressed concisely below:
\begin{equation}
 \label{biLag}
  \alpha'^2\mathcal{L}^{\text{BI}} = 1-\sqrt{1+\frac{(\alpha')^2}{2} (F_{\mu\nu}F^{\mu\nu})-\frac{(\alpha')^4}{16}(F_{\mu\nu}\tilde{F}^{\mu\nu})^2}\,.
\end{equation}
Throughout the remainder of the paper, we will set $\alpha'=1$ in the amplitude expressions.

\dj{Prolly add at least fermionic actions so maybe N=1 in 10D, and mention dimensional reduction to get maximally susy in lower D?}
\begin{equation}
 \label{vaLag}
  \alpha'^2\mathcal{L}^{\text{VA}} =\ldots
  \,.
\end{equation}

\begin{equation}
 \label{dbivaLag}
    \alpha'^2\mathcal{L}^{\text{DBIVA}} =\ldots
  \,.
\end{equation}

\dj{Explain that tree-level at least DBIVA amplitudes are  a double copy between max sym YM and pions, then maybe go into abelianized open string story. }

   DBIVA appears naturally in the field theory limit of the \textit{abelianized open superstring} \cite{Green:1982sw}. Tree-level amplitudes for the open superstring (OSS) ~\cite{Mafra:2011nv,Mafra:2011nw}  can be constructed from a double copy \cite{Broedel:2013tta} between maximally supersymmetric super-Yang-Mills (sYM) and $Z$-theory amplitudes~\cite{Carrasco:2016ldy,Carrasco:2016ygv,Mafra:2016mcc}, as follows,
\begin{equation}
A^{\text{OSS}}= Z \otimes A^{ \text{sYM}}
\end{equation}
where $\otimes$ implements the field theory double copy \cite{KLT, BCJ}. To recover the abelian sector, one simply sums over all color orderings on the side of the bi-colored $Z$-theory amplitudes that obey string monodromy relations. In the field theory limit, the observables for so-called \textit{abelianized} $Z$-theory are simply the amplitudes generated by the NLSM Lagrangian \cite{Carrasco:2016ldy}. That is, given a $n$-point amplitude in $Z$-theory, with field theory ordering $a=(a_1,a_2,...,a_n)$ and string theoretic color ordering $A=(A_1,A_2,...,A_n)$, one finds the following relation:
\begin{equation}
A^{\text{NLSM}}{(a_1,a_2,...,a_n)} \equiv \lim_{\alpha' \to 0} (\alpha')^{2-n}  Z_{\times}{(a_1,a_2,...,a_n)}
\end{equation}
where $Z_{\times}$ are abelianized $Z$-theory amplitudes of \cite{Carrasco:2016ldy}, defined by summing all possible orderings of Chan-Paton factors:
\begin{equation}
Z_{\times}(a_1,a_2,...,a_n)\equiv \sum_{A \in S^{n-1}} Z_{(A_1,...,A_{n-1},n)}{(a_1,a_2,...,a_n)}
\end{equation}
Since the field theory limit of abelian $Z$-theory produces NLSM amplitudes, the field theory limit of the abelianized OSS, gives rise to DBIVA observables at leading order:
\begin{equation}
\boxed
{\begin{aligned} 
\lim_{\alpha' \to 0 }Z^{\text{OSS}}_{\times}&= A^{\text{NLSM}} \otimes A^{\text{BAS}}\equiv A^{\text{NLSM}}
\\
\Rightarrow \quad\lim_{\alpha' \to 0 }A^{\text{OSS}}_{\times}&= A^{\text{NLSM}} \otimes A^{ \text{sYM}}\equiv A^{\text{DBIVA}}
\end{aligned}}
\end{equation}
These two classes of amplitudes, NLSM and DBIVA, are the observables that we will consider in the multi-loop studies of this work. Additional details on the bi-colored $Z$-theory amplitudes these string-theoretic double-copy constructions can be found in \cite{Broedel:2013tta,Carrasco:2016ldy,Carrasco:2016ygv,Mafra:2016mcc,Azevedo:2018dgo} and references therein.

By far the most important feature of these amplitudes relevant for our analysis is that they are generated by purely \textit{even-point} (EP) contacts. That is, the effective actions of EP theories are consistent with the following interaction Lagrangian:
\begin{equation}
\mathcal{L}^{\text{abelian}}_{\text{int.}} = \sum_{k=2} \alpha'^{n-2} \mathcal{O}^{(2n)}_{2k} 
\end{equation}
where $2k$ counts the field content of the effective operator, and $2n$ counts the mass-dimension determined by the number of fields and derivatives contained in the operator. The dimensionful scale, $\alpha'$, carries mass dimension $[\alpha']=-2$ in $D=4$.

The even-point nature of this class of theories make them comparatively easy to compute relative to their arbitrary-multiplicity cousins. Indeed, the observation that will drive many of the multi-loop results is the following: \textit{amplitudes for EP theories at next-to-next-to-leading-order (NNLO) in $\alpha'$ are recursively one-loop}. To gain some familiarity with why this is the case, consider the interactions needed to construct two loop amplitudes for a simple $\varphi^{2k}$-scalar theory:
\begin{equation}\label{eq:evenPointL}
\mathcal{L}_{\varphi^{2k}} = \frac{1}{2}(\partial \varphi)^2 + c_4 \varphi^4 + c_6 \varphi^6+ c_8 \varphi^8+\cdots 
\end{equation}
While the Feynman rules for this theory are trivial, unlike those needed for the vectors and fermions in the DBIVA supermultiplet and the derivatively coupled pions in NLSM, the Feynman diagrams needed to compute amplitudes are identical to those of $A^{\text{OSS}}_{\times}$. Indeed, the two-loop amplitude for $\varphi^{2k}$ theory is generated simply by evaluating the following set of scalar 1PI integrals:

@@

\dj{A few words about higher derivative operators.  Point out that HD operators of DBIVA from abelianized open string do *not* satisfy U1.}

\subsection{On-shell Unitarity methods }\label{sec:genU}
\dj{on-shell unitarity methods: unitarity based integrand verification and construction (maybe like a page), 4D and D-dim completeness relations. }

To perform the loop-level calculations in this paper, we will construct the integrand in $D$-dimensions using generalized unitarity cuts \cite{Bern:1995db,Bern:1996ja}. In the kinematic regime where the internal loop momenta go on-shell, the residue is a product of on-shell tree level diagrams, summed over all physical states that can cross the cut. For example, the kinematic numerator of the one-loop integrand $\mathcal{I}_{\text{1-loop}}$ of pure Born-Infeld theory should satisfy the following constraint:
\begin{equation}
\label{GUCut}
\text{Cut}( \mathcal{I}_{\text{1-loop}}) = \sum_{h_i\in \text{states}} \mathcal{M}^{\text{BI}}_4(l_1^{h_1},-l_2^{\bar{h}_2})\mathcal{M}^{\text{BI}}_4(l_2^{h_2},-l_1^{\bar{h}_1})\,,
\end{equation}  
where $\text{Cut}( \mathcal{I}_{\text{1-loop}}) $ extracts the kinematic numerator by imposing a maximal cut on the one-loop integrand 
\begin{equation}
\text{Cut}( \mathcal{I}_{\text{1-loop}})  \equiv l_1^2l_2^2(\mathcal{I}_{\text{1-loop}})\Big|_{l_1^2,l_2^2\rightarrow 0}\,.
\end{equation}
To carry out the sum over states in \eqn{GUCut} we apply the following $D$-dimensional completeness relation for on-shell polarizations,
\begin{equation}\label{stateSum}
\sum_{\text{states}}\varepsilon^\mu_{(l)}\varepsilon^\nu_{(-l)} = \eta^{\mu\nu} - \frac{l^\mu q^\nu+l^\nu q^\mu}{l\cdot q},
\end{equation}
with null reference momentum, $q^2=0$. The dependence of $q$ is a gauge choice that disappears when on-shell kinematic constraints are imposed. At loop level, the state-sum of \eqn{stateSum} can potentially lead to dimension dependence in the $D$-dimensional integrand. This is due to terms that contain products of internal polarizations:
\begin{equation}
\begin{aligned}
 \mathcal{I}_{\text{1-loop}} &\supset \sum_{ \text{states}} \varepsilon^\mu_{(l)}\varepsilon^\nu_{(-l)} \eta_{\mu\nu}
 \\
 &=\left(\eta^{\mu\nu} - \frac{l^\mu q^\nu+l^\nu q^\mu}{l\cdot q}\right) \eta_{\mu\nu}
 \\
 &=D_s-2
 \end{aligned}
\end{equation}
 where $D_s$ is the spin-dimension of the internal photons, and where $D_s-2$ just counts the number of photon states. To calculate the loop-level amplitudes in this work, we use dimensional regularization to regulate the integrals. In this work we adopt the conventions of  \cite{collins_1984,Bern:2002zk} and take the spin-dimension, $D_s$, to be
the same as the dimension appearing in dimensional regularization, $D=4-2\epsilon$. We will use them interchangeably throughout. As shown in \eqn{eq:2loopEven} of the introduction, many of the diagrams that contribute to EP loop-level amplitudes are scaleless, and thus vanish in dimensional regularization.

\subsection{One-loop Unitarity}
\dj{2.6 tensor reduction, relevant 1-loop basis integrals, Cut Extraction of One-loop Amplitude Integral Coefficients}

It is well known that one-loop amplitudes are spanned by a small basis of scalar integrals \cite{Forde:2007mi,Badger:2008cm,ElvangHuangReview}. This property can be exposed via a $D$-dimensional integral reduction algorithm due to Passarino and Veltman \cite{Passarino:1978jh}.  Darren Forde showed blah blah.

\section{Even-Point Multiloop Unitarity}
\dj{Emus are soft-feathered, brown, flightless birds with long necks and legs, and can reach up to 1.9 metres (6.2 ft) in height. Emus can travel great distances, and when necessary can sprint at 48 km/h (30 mph). They forage for a variety of plants and insects, but have been known to go for weeks without eating. They drink infrequently, but take in copious amounts of water when the opportunity arises.}

At two-loops we can exploit the even-point properties of the theories of interest in two ways -- first in the integral construction via unitarity.  The second via tensor reduction of these special basis integrals.

\subsection{Recursive integrals}
\label{sec:recInt}
With the scaleless behavior of two-loop integrals in mind, the only integrals that contribute at two-loop order are given in \Fig{fig:contrib2Loop}. As one can see, only the four-point tree-level diagrams are required for cut construction for the pure Born-Infeld (BI) part of the amplitude through the two-loop. As such, the physical part of the two-loop integrand is completely fixed by a set of cuts down to the four-point $\mathcal{M}^{\text{BI}}_4=t_8F^4$ vertex:
\begin{equation}
\text{Cut}(\mathcal{I}_{\text{2-loop}} )= \sum_{\text{states}} \mathcal{M}^{\text{BI}}_4\mathcal{M}^{\text{BI}}_4\mathcal{M}^{\text{BI}}_4
\end{equation}
Similar to \eqn{GUCut}, $\text{Cut}(\mathcal{I})$ extracts the kinematic numerator for the cut configurations captured at two-loop in \Fig{fig:contrib2Loop}.
